### 2017-03-08 real/draft.tex
[draft] Major clean up
* Remove most comments
* remove todo's at beginning
* little fixes
* add plot of tardis histogram

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{170221}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}

@@

\newcommand{\refeq}[1]{Eq.~(\ref{#1})}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\refsec}[1]{Sec.~\ref{sec:#1}}
\DeclareMathOperator{\erfc}{erfc}
\DeclareMathOperator{\Expect}{\mathbb{E}}
\newcommand{\expect}[1]{\Expect\left[#1\right]}
\newcommand{\expectest}[1]{\widehat{\Expect\left[#1\right]}}
\DeclareMathOperator{\GammaDist}{Gamma}
\DeclareMathOperator{\GaussianDist}{\mathcal{N}}
\DeclareMathOperator{\InvGammaDist}{InvGamma}
\newcommand{\Kalpha}{{K_\alpha}}
\newcommand{\Kbeta}{{K_\beta}}
\newcommand{\npack}{{n_p}}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{{L_{\rm max}}}
\newcommand{\Lumtot}{Q}
\newcommand{\lumtot}{q}
\newcommand{\Lum}{L}
\newcommand{\lum}{\ell}
% \newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\rmdx}[1]{\dd{#1}} % differential
\newcommand{\firstDeriv}[1]{\frac{\partial}{\partial #1}}
\newcommand{\secDeriv}[1]{\frac{\partial^2}{\partial #1^2}} % second partial derivative
\newcommand{\secPartial}[2]{\frac{\partial^2}{\partial #1 \, \partial #2}} % second partial derivative
\newcommand{\tardis}{TARDIS}
\DeclareMathOperator{\Variance}{\mathbb{V}}
\newcommand{\variance}[1]{\Variance\left[#1\right]}

@@

\section{What's left to do}

\subsection{why asymptotic approx. has larger variance}
\begin{enumerate}
  \item For $\alpha=1.5, \beta=60, n=14000$ still get 50 \% larger uncertainty from asymptotic compared to summation over $N$. Is it due very asymmetric one-packet distribution? Try larger $\alpha$, for example $\alpha = 30$.
  \item  Another option: $\int \rmdx{\alpha} \rmdx{\beta}$, use true values in simulation. Where does uncertainty come from? From $\lambda$ or $\mu, \sigma^2$?
  \item Are finite number of terms in $\sum_N$ a problem? Check

\subsection{Plots}

\begin{enumerate}
  \item generate a toy spectrum, take frequencies from actual tardis,
  generate luminosities from a $\GammaDist(\alpha, \beta)$, no $\nu$
  dependence of $\alpha, \beta$ for clarity , plot spectrum with
  uncertainty bands, show it early in the paper, then describe where
  it comes from
  \item compare asymptotic and $\sum_N$ $p(\Lumtot \cond \bmy)$ on a
  plot, add what happens if only $\sqrt{n}$ uncertainty is used: that
  would neglect uncertainty on the single-packet distribution, show
  case when parameters determined from a single bin with few samples,
  should make large difference. Fix $\lambda$, what are the uncertainties now?
  \item Statistics of replicas: again compare \refeq{spe} to \refeq{ral}. Teach a lesson: from a single data set, prediction will be centered there
\end{enumerate}

\subsection{Other issues that need to be resolved}

\begin{enumerate}

\item Clean up integrals using \verb=\rmdx= rather than $d$.

\item Write everything in terms of frequency $\nu$ or wavelength
  $\lambda$

\item Now that we have removed direct analysis of \tardis\ from the
  paper, there is no need to define an overall parameterization of the
  Gamma parameters $\alpha(\nu_b) = \sum_k a_k \nu_b^k$ and
  $\beta(\nu_b) = \sum_k b_k \nu_b^k$. However, the possibility of
  linking the analyses of all the bins by this and similar methods
  should be highlighted. Perhaps in the Conclusions?

\item \textbf{\tardis} move tardis section into separate document in
  this git repo. Better to keep it out of the paper. Cite tardis when
  showing the toy spectrum. Mention that in reality $\alpha$ and
  $\beta$ are $\nu$ dependent

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{spectrum_bin}
  \includegraphics[width=0.48\textwidth]{spectrum_all}
  \caption{Spectrum with smallest 2 $\sigma$ credibility interval. Left: only packets from bin contribute. Right: a}
\label{fig:spectrum}
\end{figure}

@@

We consider a single computer simulation which generates a luminosity
spectrum from $\npack$ photon packets, each having a particular
frequency $\nu_i$ and luminosity $\lum_i$. Together, these frequencies
and luminosities form the raw simulation data
$\bmy = \{ (\nu_i,\lum_i): i = 1,\ldots,\npack\}$. Typically, the
frequency spectrum is partitioned into bins of width $\Delta\nu$
centered on frequencies $\nu_b$ resulting in bin intervals
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$.  Let $n_b$ be the number
of packets having a frequency falling into a particular bin $b$, and
let the vector $\bml_b = \{(\ell_{b,j}): j = 1,\ldots,n_b\}$ be the
set of luminosities of those packets. The data set is correspondingly
cast into the more compact histogram form
\begin{align}
  \lleq{sdb}
  \bmy = \{(n_b,\bml_b): b = 1,\ldots,B\}
\end{align}
with loss of information limited to replacing the exact packet
frequencies $\nu_i$ by the approximate bin frequency $\nu_b$. The
total luminosity in bin $b$ is
\begin{align}
  \lleq{sdc}
  \lumtot_b
  &= \sum_{j=1}^{n_b} \lum_{b,j}.
\end{align}
Within a given model $\hypo$, the data $\bmy$ are only one of many
possible realisations of the corresponding random variables; for
example, starting the simulation with a different seed would result in
different output data. We denote the general random variables by the
corresponding upper-case symbols,
\begin{align}
  \lleq{sdd}
  \bmL_b &= \{(\Lum_{b,j}): j = 1,\ldots,N_b\}\\
  \lleq{sde}
  \bmY &= \{(N_b, \bmL_b): b = 1,\ldots,B\},\\
  \lleq{sdf}
  \Lumtot_b &= \sum_{j=1}^{N_b} \Lum_{b,j},
\end{align}
Lower-case symbols such as $\bmy, n_b,\bml_b$ etc.\ are strictly
reserved for the simulated data at hand.

@@

Various physical processes such as absorption result in luminosity
spectra with strong features. The model we develop below takes into
account the variability introduced by these features in a two-step
process termed a \textit{compound process} in the statistics
literature.
%
In the first step, the number of packets $N$ is modelled by means of a
separate Poisson distribution $p(N\cond\lambda)$ for each bin, where
the change in magnitude of $\lambda$ from bin to bin will capture most
of the variability of the spectrum. Given $N$, the single-packet
luminosity distribution is modelled in the second step by means of a
packet likelihood for $\Lum_{b,j}$ with packet parameters, a packet
prior and, of course, a packet parameter posterior. The single-packet
distribution contains the crucial information which allows us to
extend previous methods to cases where the number of packets in a bin
is small.
%%%%%%%%%%%%%%%%%%%
\fred{Refs: Zech, maybe others, too}

@@

First, we eliminate the parameter $\lambda$ of the Poisson
multiplicity distribution for the number of packets $N$,
\begin{align}
  \lleq{pmb}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^{N}}{N!}
  \quad N = 0,1,\ldots,\infty.
\end{align}
The simulation number of packets $n$ is naturally presumed to have
arisen from the same Poisson distribution with the same $\lambda$.
The posterior for $\lambda$ is found from the data $n$ via Bayes'
Theorem,
\begin{align}
  \lleq{pmc}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
We use a prior
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a}
\end{align}
which is ``improper'', meaning that
$\int_0^\infty \rmdx{\lambda}\,p(\lambda)$ is known only up to a
constant $c$. An improper prior is not a problem because $c$ cancels
out when calculating the posterior, provided that $n \ge 1$; i.e., at
least one packet is observed in the bin.\footnote{For a uniform prior,
  the case $a{=}0$, we require of course a parameter
  $\lambda_{\rm max}$. As long as $\lambda_{\rm max}$ is large enough,
  it would also cancel.}, The posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n)
  &= \GammaDist(\lambda \cond \alpha = n{-}a{+}1 \,,\, \beta = 1)\,,
\end{align}
where the $\GammaDist$ distribution with shape parameter $\alpha > 0$ and
rate parameter $\beta > 0$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x\cond \alpha, \beta)
  &= \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\,,
  \quad 0 \leq x < \infty.
\end{align}
The evidence is
$p(n) =  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda) = c\,(n-a)!/n!$
% \begin{align}
%   \lleq{pmg}
%   p(n)
%   &=  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda)
%     \ =\ c\frac{(n-a)!}{n!} \,, \\
% \end{align}
and inserting Eqs.~(\ref{pmb}) and (\ref{pmc}), the prediction for $N$
given $n$ is a negative binomial distribution
\begin{align}
  \lleq{pmh}
  p(N\cond n) %
  &= \int_0^\infty \rmdx{\lambda}\, p(N\cond\lambda)\,p(\lambda\cond n)
  \ =\ \binom{N{+}n{-}a}{N} \left(\frac{1}{2}\right)^{N+n-a+1}\, .
    %\ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
\end{align}
A hyperparameter choice $a=0$ yields the uniform prior, $a=1/2$ gives
the Jeffreys and reference prior, and $a=1$ is the
transformation-group result advocated by
Jaynes~\cite[Ch. 12]{jaynes2003probability}. The uniform prior has no
solid information-theoretic foundation in this case reducing the
choice to $a \in \{1/2, 1\}$. We examine the consequence for the
extreme case that no packet is observed: $P(N=0 \cond n=0) =
2^{a-1}$. For $a=1$, any $N>0$ is excluded which is undesirable unless
we accept that repeating the simulation will never yield a packet in
the current bin. Hence we are led to Jeffreys prior, $a=1/2$.

@@

Next, we consider detailed modelling of $p(\bmL\cond N,\bmphi)$ which
was written in Eq.~(\ref{slk}) as a product of single-packet
distributions $p(L_j\cond \bmphi)$, meaning that single packets are
assumed to be mutually independent and to follow the same distribution
independent of $N$.
% 
Which specific probability function to choose for $p(L_j\cond\bmphi)$
depends, of course, on the specific simulation data. In this section,
we choose it to be the same Gamma distribution for all $L_j$
with parameters $\bmphi = (\alpha,\beta)$,
\begin{align}
  \lleq{spc}
  p(L_j\cond \bmphi) &= \GammaDist(L_j\cond \alpha,\beta)
\end{align}
which has the nice property that the convolution (\ref{slm}) results
in another Gamma distribution,
\begin{align}
  \lleq{spd}
  \int \rmdx{\bmL}\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \GammaDist(Q\cond N\alpha, \beta),
\end{align}
and so Eq.~(\ref{gmh}) becomes in this example
\begin{align}
  \lleq{spe}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    p(\alpha,\beta\cond n,\bml).
\end{align}
The posterior for the parameters is calculated by yet another
application of Bayes' Theorem,
\begin{align}
  \lleq{spf}
  p(\alpha,\beta\cond n,\bml)
  &= \frac{p(n,\bml\cond \alpha,\beta)\,p(\alpha,\beta)}{p(n,\bml)}
\end{align}
where the data $(n,\bml)$ must naturally follow the same
Gamma likelihood as $(N,\bmL)$, i.e.\
\begin{align}
  \lleq{spg}
  p(n,\bml\cond \alpha,\beta)
  & = \prod_{j=1}^n \GammaDist(\ell_j\cond\alpha,\beta)
    \ =\ \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n}
\end{align}
with the data summarized by the sufficient statistics $n$,
$q \equiv \sum_{j=1}^n \ell_j$, and $r \equiv \prod_{j=1}^n \ell_j$.

\fred{Here and above, $n$ does not have to be the same in
  $P(N \cond n)$ and $p(n,\bml\cond \alpha,\beta)$. We can use packets
  from other bins if we assume $\alpha, \beta$ are the same in those bins.}
\hans{Please explain more fully; not clear.}

It only remains to explicitly evaluate the prior $p(\alpha,\beta)$
appearing in \refeq{spf}.
%
The authors of \cite{moala2013bayesian} present an overview of
possible choices indicating that their \emph{modified MDIP} prior
performed best in the few examples considered there. We do not follow
their recommendation because their modification is not well motivated
and it seems to accidentally perform better because of the particular
examples considered. The other conclusion they have is that the
Jeffreys, reference, and other noninformative priors with similarly
solid motivation have comparable performance. They all exhibit the
standard form for a scale or rate parameter $\propto 1/\beta$ and only
differ in the $\alpha$ dependence. We disregard the reference prior as
it forces us to identify either $\alpha$ or $\beta$ as \emph{the}
nuisance parameter but we consider \emph{both} as nuisance
parameters. Hence we select the Jeffreys prior
\hans{What is $\Psi$?}
\begin{align}
  \lleq{sph}
  p(\alpha,\beta)
  &=  C \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta} \, .
  % \frac{1}{(\alpha_{\rm max}{-}1)\beta_{\rm max}}
  % \qquad 1 < \alpha < \alpha_{\rm max},\; 0 < \beta < \beta_{\rm max}.
\end{align}
% since $\alpha_{\rm max}$ and $\beta_{\rm max}$ cancel in (\ref{spf}).
With these choices, the posterior normalization is %
\hans{``Posterior normalization'' = Evidence? With improper priors,
  should the upper limits not be $\infty$ below?}
\begin{align}
  \lleq{spj}
  p(n, \bml) &= C \int_0^{\alpha_{\rm max}} d\alpha \int_0^{\beta_{\rm max}} d\beta\;
  \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n},
\end{align}
where the upper integration limits
$(\alpha_{\rm max}, \beta_{\rm max})$ should formally extend to
$\infty$ but in practice have to be chosen by trial and error to be
large enough to contain the relevant contribution to the integral.
The prior's constant of proportionality $C$ cancels in the final
result and we may set $C\equiv 1$ to obtain
\begin{align}
  \lleq{spk}
  p(Q\cond n,\bml)
  &= \frac{1}{p(n, \bml)}
    \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n} \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta}.
\end{align}

\fred{I removed a lot of cruft and swapped the symbols
  $x \to \ell, \ell \to \ell^\prime$ so it's in line with the previous section}

\hans{This and the next section interrupt the actual calculation of
  exact results in Sections
  \ref{sec:multiplicity}--\ref{sec:single-packet} and the asymptotic
  ones in Section \ref{sec:asymptotic}, so I would suggest moving them
  to the end just before the conclusions. }

\tardis{} can output two kinds of packets, real and virtual. Here we
focus on the real packets and denote the raw output of \tardis{} by
primed symbols, for example for the $i$-{th} packet
$\qty(\nu_i^\prime, \ell_i^\prime)$.  A histogram plot
\fredmargin{include plot} of the luminosities from a \tardis{} run
indicates a strongly skewed distribution with its single mode close to
the maximum luminosity $\lmax^\prime$, while there is a long tail for
small $\ell$. These features can be accommodated by a $\GammaDist$
distribution as in \refsec{single-packet} if we transform the
luminosities to
\begin{align}
  \lleq{gmf}
  \ell_i \equiv 1- \frac{\ell_i^\prime}{\lmax^\prime} > 0
\end{align}
and similarly $L_i \equiv 1 - (L_i^\prime/\lmax^\prime)$. We assume
the total number of samples across all bins is large, so that the effect
of considering $\lmax^\prime$ a constant independent of the data is
negligible. We checked that leaving $\lmax^\prime$ as a free parameter
in a fit gave equivalent results.
%
Numerically, we take the largest luminosity of all \tardis{} packets
across the entire frequency range and increase it slightly to avoid
numerical difficulties at the end point
\begin{align}
  \lleq{gme}
  \lmax^\prime \equiv (1+\epsilon) \times \max_i \ell_i^\prime,
\end{align}
where $ \epsilon=10^{-6}$. We carry out all calculations with
$p(\Lumtot \cond n,\bml)$ and only in the end transform back to a
probability density in the raw coordinates via
\begin{align}
  \lleq{gmg}
  p(\Lumtot^\prime  \cond n, \bml^\prime)
  &= p(\Lumtot  \cond n, \bml) \frac{1}{\lmax^\prime} \,.
\end{align}
While we also rescale the frequencies by a constant from
$\mathcal{O}(10^{15})$ to $\mathcal{O}(1)$, the most important
numerical simplification arises because of the transformation of the
luminosities. Rewriting \refeq{sll} formally as
\begin{align}
  \lleq{gmhb}
  p(Q  \cond n,\bml)
  & = \sum_{N} p(N \cond n)\; p(Q \cond N, n, \bml),
\end{align}
we see that the right-hand side is a mixture density. In the \tardis{}
application, the distribution of a single-packet luminosity,
$p\qty(\ell_i' \cond \bmphi)$ is narrow in the sense that the standard
deviation relative to the mean is at the percent level. This implies
that each $p(Q^\prime \cond N, n, \bml^\prime)$ contributes one narrow
peak and all peaks are clearly separated in the superposition
$ p(Q^\prime \cond n,\bml^\prime)$. The transformation
$\ell^\prime \to \ell$ effects a strong overlap between the
distributions $p(\Lumtot \cond N, n, \bml)$ with similar $N$, so
$p(\Lumtot \cond n, \bml)$ can be well approximated by a unimodal
distribution and thus $\Lumtot$ rather than $\Lumtot^\prime$ is the
preferred coordinate basis.
%
\fredmargin{show hedgehog plot}

The parameters $\qty(\alpha, \beta)$ of the single-packet distribution
$\GammaDist(\ell_i \cond \alpha, \beta)$ may, of course, depend on the
frequency. % $\nu_i$
This effect could be modelled, for example, by a functional
parametrization $\alpha = \alpha(\nu \cond \bmphi)$ and
$\beta = \beta(\nu \cond \bmphi)$ in dependence on the frequency $\nu$
and on hyperparameters $\bmphi$ such that integrals over
$\alpha, \beta$ would turn into integrals over $\bmphi$. For the
sake of clarity, let rather us consider only the simplest case where
$\alpha, \beta$ are the same in every bin such that \refeq{spk}
becomes
\begin{align}
  \lleq{gmi}
  p(Q\cond \npack,\bmy)
  &=     \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    p(\alpha, \beta \cond \npack, \bmy) \, .
\end{align}
In words, the posterior $ p(\alpha, \beta \cond \npack, \bmy)$ uses
all packets to essentially fix $\alpha, \beta$ if $\npack \gg 1$ and
allows meaningful predictions of $\Lumtot$ even in bins where
\emph{no} packet is observed; i.e., $n=0$. The only binning left is
performed to count $n$ the number of packets that fall into a given
frequency bin. This strategy is applicable, for example, when
absorption or emission lines in the spectrum abruptly alter the number
of packets for nearby frequencies but the luminosities are unaffected
by those line features. Note that if the dependence on $\bmphi$ can be
modeled with few parameters, the total number of parameters may be
drastically reduced: instead of determining $\alpha, \beta$ separately for
each of a few thousand bins, there are potentially only a handful of
elements in $\bmphi$ for the entire spectrum. In that case, the
integral in \refeq{gmi} does not depend on the bin but only on $N$ so
its value may be cached to speed up the simultaneous evaluation across
all bins.

>>>>

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{170308}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}

@@

\newcommand{\refeq}[1]{Eq.~(\ref{#1})}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\refsec}[1]{Sec.~\ref{sec:#1}}
\DeclareMathOperator{\erfc}{erfc}
\DeclareMathOperator{\Expect}{\mathbb{E}}
\newcommand{\expect}[1]{\Expect\left[#1\right]}
\newcommand{\expectest}[1]{\widehat{\Expect\left[#1\right]}}
\DeclareMathOperator{\GammaDist}{Gamma}
\DeclareMathOperator{\GaussianDist}{\mathcal{N}}
\DeclareMathOperator{\InvGammaDist}{InvGamma}
\newcommand{\Kalpha}{{K_\alpha}}
\newcommand{\Kbeta}{{K_\beta}}
\newcommand{\npack}{{n_p}}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{{L_{\rm max}}}
\newcommand{\Lumtot}{Q}
\newcommand{\lumtot}{q}
\newcommand{\Lum}{L}
\newcommand{\lum}{\ell}
\newcommand{\rmdx}[1]{\dd{#1}} % differential
\newcommand{\firstDeriv}[1]{\frac{\partial}{\partial #1}}
\newcommand{\secDeriv}[1]{\frac{\partial^2}{\partial #1^2}} % second partial derivative
\newcommand{\secPartial}[2]{\frac{\partial^2}{\partial #1 \, \partial #2}} % second partial derivative
\newcommand{\tardis}{TARDIS}
\DeclareMathOperator{\Variance}{\mathbb{V}}
\newcommand{\variance}[1]{\Variance\left[#1\right]}

@@

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{spectrum_bin}
  \includegraphics[width=0.48\textwidth]{spectrum_all}
  \caption{Spectrum with most likely value of $Q$ (dot) and smallest
    2$\sigma$ credibility interval (blue band). Left: only packets
    from single bin contribute to the single-packet luminosity
    distribution in that bin. If $n<8$ packets observed, no reliable
    prediction is available and bins are masked out (red band) Right:
    Same packets as on the left but the single-packet luminosity
    distribution is inferred from all packets. }
\label{fig:spectrum}
\end{figure}

@@

We consider a single computer simulation which generates a luminosity
spectrum from $\npack$ photon packets, each having a particular
frequency $\nu_i$ and luminosity $\lum_i$. Together, these frequencies
and luminosities form the raw simulation data
$\bmy = \{ (\nu_i,\lum_i): i = 1,\ldots,\npack\}$. Typically, the
frequency spectrum is partitioned into bins of width $\Delta\nu$
centered on frequencies $\nu_b$ resulting in bin intervals
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$.  Let $n_b$ be the number
of packets having a frequency falling into a particular bin $b$, and
let the vector $\bml_b = \{(\ell_{b,j}): j = 1,\ldots,n_b\}$ be the
set of luminosities of those packets. The data set is correspondingly
cast into the more compact histogram form
\begin{align}
  \lleq{sdb}
  \bmy = \{(n_b,\bml_b): b = 1,\ldots,B\}
\end{align}
with loss of information limited to replacing the exact packet
frequencies $\nu_i$ by the approximate bin frequency $\nu_b$. The
total luminosity in bin $b$ is
\begin{align}
  \lleq{sdc}
  \lumtot_b
  &= \sum_{j=1}^{n_b} \lum_{b,j}.
\end{align}
Within a given model $\hypo$, the data $\bmy$ are only one of many
possible realisations of the corresponding random variables; for
example, starting the simulation with a different seed would result in
different output data. Following standard notation, we denote the
general random variables by the corresponding upper-case symbols,
\begin{align}
  \lleq{sdd}
  \bmL_b &= \{(\Lum_{b,j}): j = 1,\ldots,N_b\}\\
  \lleq{sde}
  \bmY &= \{(N_b, \bmL_b): b = 1,\ldots,B\},\\
  \lleq{sdf}
  \Lumtot_b &= \sum_{j=1}^{N_b} \Lum_{b,j}.
\end{align}
Lower-case symbols such as $\bmy, n_b,\bml_b$ etc.\ are strictly
reserved for the realisations; i.e., the simulated data at hand.

@@

Various physical processes such as absorption result in luminosity
spectra with strong features. The model we develop below takes into
account the variability introduced by these features in a two-step
process termed a \textit{compound process} in the statistics
literature.
%
In the first step, the number of packets $N$ is modelled by a Poisson
distribution $p(N\cond\lambda)$, separately for each bin, where the
change in magnitude of $\lambda$ from bin to bin will capture most of
the variability of the spectrum. Given $N$, the single-packet
luminosity distribution is modelled in the second step by means of a
packet likelihood for $\Lum_{b,j}$ with packet parameters, a packet
prior and, of course, a packet parameter posterior. The single-packet
distribution contains the crucial information which allows us to
extend previous
methods~\cite{bohm_comparison_2012,bohm_statistics_2014} to cases
where the number of packets in a bin is small and even zero.

@@

First, we eliminate the parameter $\lambda$ of the Poisson
multiplicity distribution for the number of packets $N$,
\begin{align}
  \lleq{pmb}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^{N}}{N!}
  \quad N = 0,1,\ldots,\infty.
\end{align}
The simulation number of packets $n$ is naturally presumed to have
arisen from the same Poisson distribution with the same $\lambda$.
The posterior for $\lambda$ is found from the data $n$ via Bayes'
Theorem,
\begin{align}
  \lleq{pmc}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
We use a prior
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a}
\end{align}
which is ``improper'', meaning that
$\int_0^\infty \rmdx{\lambda}\,p(\lambda)$ is known only up to a
constant $c$.\footnote{For a uniform prior, the case $a{=}0$, we
  additionally require an upper limit $\lambda_{\rm max}$. As long as
  $\lambda_{\rm max}$ is large enough, it would also cancel.}, The
posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n)
  &= \GammaDist(\lambda \cond \alpha = n{-}a{+}1 \,,\, \beta = 1)\,,
\end{align}
where the $\GammaDist$ distribution with shape parameter $\alpha > 0$ and
rate parameter $\beta > 0$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x\cond \alpha, \beta)
  &= \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\,,
  \quad 0 \leq x < \infty.
\end{align}
The evidence is
$p(n) =  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda) = c\,(n-a)!/n!$.
% \begin{align}
%   \lleq{pmg}
%   p(n)
%   &=  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda)
%     \ =\ c\frac{(n-a)!}{n!} \,, \\
% \end{align}
Upon inserting Eqs.~(\ref{pmb}) and (\ref{pmc}), $c$ cancels out and
the integral over $\lambda$ can be performed analytically resulting in
a negative binomial distribution that is proper for any $n$
\begin{align}
  \lleq{pmh}
  p(N\cond n) %
  &= \int_0^\infty \rmdx{\lambda}\, p(N\cond\lambda)\,p(\lambda\cond n)
  \ =\ \binom{N{+}n{-}a}{N} \left(\frac{1}{2}\right)^{N+n-a+1}\, .
    %\ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
\end{align}
A hyperparameter choice $a=0$ yields the uniform prior, $a=1/2$ gives
the Jeffreys and reference prior, and $a=1$ is the
transformation-group result advocated by
Jaynes~\cite[Ch. 12]{jaynes2003probability}. To decide on $a$, we note
that the uniform prior has no solid information-theoretic foundation
in this case reducing the choice to $a \in \{1/2, 1\}$. We examine the
consequence for the extreme case that no packet is observed:
$P(N=0 \cond n=0) = 2^{a-1}$. For $a=1$, any $N>0$ is excluded which
is undesirable unless we accept that repeating the simulation will
never yield a packet in the current bin. Hence we settle on Jeffreys
prior, $a=1/2$, for any numerical evaluation carried out below.

@@

Next, we consider detailed modelling of $p(\bmL\cond N,\bmphi)$ which
was written in Eq.~(\ref{slk}) as a product of single-packet
distributions $p(L_j\cond \bmphi)$, meaning that single packets are
assumed to be mutually independent and to follow the same distribution
independent of $N$ and the exact frequency within the bin under consideration.
%
Which specific probability function to choose for $p(L_j\cond\bmphi)$
depends, of course, on the specific simulation data. In this section,
we choose it to be the same Gamma distribution for all $L_j$
with parameters $\bmphi = (\alpha,\beta)$,
\begin{align}
  \lleq{spc}
  p(L_j\cond \bmphi) &= \GammaDist(L_j\cond \alpha,\beta)
\end{align}
which has the nice property that the convolution \refeq{slm} for
$N > 0$ results in another Gamma distribution,
\begin{align}
  \lleq{spd}
  \int \rmdx{\bmL}\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \GammaDist(Q\cond N\alpha, \beta).
\end{align}
For the special case $N=0$, there is no luminosity because there is no
packet, so we define
\begin{align}
  \lleq{spl}
   p(Q\cond N, \alpha, \beta) &\equiv
   \begin{cases}
     \delta(Q),     &N=0,\\
     \GammaDist(Q\cond N\alpha, \beta), & N > 0 \, ,
   \end{cases}
 \end{align}
 with the Dirac $\delta$ function and so \refeq{gmh} simplifies to
\begin{align}
  \lleq{spe}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    p(Q\cond N, \alpha, \beta)\;
    % \GammaDist(Q\cond N\alpha, \beta)\;
    p(\alpha,\beta\cond n,\bml).
\end{align}
The posterior for the parameters is calculated by yet another
application of Bayes' Theorem,
\begin{align}
  \lleq{spf}
  p(\alpha,\beta\cond n,\bml)
  &= \frac{p(n,\bml\cond \alpha,\beta)\,p(\alpha,\beta)}{p(n,\bml)}
\end{align}
where the data $(n,\bml)$ must naturally follow the same
Gamma likelihood as $(N,\bmL)$, i.e.\
\begin{align}
  \lleq{spg}
  p(n,\bml\cond \alpha,\beta)
  & = \prod_{j=1}^n \GammaDist(\ell_j\cond\alpha,\beta)
    \ =\ \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n}
\end{align}
with the data summarized by the sufficient statistics $n$,
$q \equiv \sum_{j=1}^n \ell_j$, and $r \equiv \prod_{j=1}^n \ell_j$.

It only remains to specify the prior $p(\alpha,\beta)$ appearing in
\refeq{spf}.
%
The authors of \cite{moala2013bayesian} present an overview of
possible choices indicating that their \emph{modified MDIP} prior
performed best in the few examples considered there. We do not follow
their recommendation because their modification is not well motivated
and it seems to only accidentally perform better because of the
particular examples they considered. The other conclusion they have is
that the Jeffreys, reference, and other noninformative priors with
similarly solid motivation have comparable performance. They all
exhibit the standard form for a scale or rate parameter
$\propto 1/\beta$ and only differ in the $\alpha$ dependence. We
disregard the reference prior as it forces us to identify either
$\alpha$ or $\beta$ as \emph{the} nuisance parameter but we consider
\emph{both} as nuisance parameters. Hence we select the Jeffreys prior
\begin{align}
  \lleq{sph}
  p(\alpha,\beta)
  &=  C \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta} \, ,
\end{align}
where $\Psi$ is the digamma function. With these choices, the
posterior is nondegenerate for $n \geq 2$ and the evidence is
\begin{align}
  \lleq{spj}
  p(n, \bml) &= C \int_0^{\infty} \rmdx{\alpha} \int_0^{\infty} \rmdx{\beta}\;
  \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n}\;
  \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta}.
\end{align}
The prior's constant of proportionality $C$ cancels in the final
result and we may set $C\equiv 1$ to obtain
\begin{align}
  \lleq{spk}
  p(Q\cond n,\bml)
  &= \frac{1}{p(n, \bml)}
    \sum_{N=1}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n} \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta}
\end{align}
for $Q > 0$ and for $Q=0$ there is the additional contribution from
$N=0$ through the $\delta$ function from \refeq{spl}.
